Seminarium TI 3700\\ Data Analysis on Twitter(working title)

### 2014-03-19 authorinstr_2.tex
update

\section{Single tweet analysis}
In this section we will discuss the different types of analysis that can be done on one single tweet. This usually is also the basis for a analysis of a larger amount of tweets or even on a stream of tweets. 
\subsection{Data enrichment}
At the very basis of the analysis of a tweet is data enrichment. Because the length of a tweet is restricted to 140 characters, the users of Twitter use language in creative ways. This way of communicating makes it sometimes, even for humans, hard to understand what a tweet is about. To get a better understanding of the meaning of a tweet, a context is needed. For humans this is done by linking the concepts from the tweet to earlier acquired knowledge. The same thing can be done by computers. Two stages can be defined in this process, first the extraction of the concepts from the tweet, after that a context can be created by linking the tweet to other sources.
\subsubsection{Concept extraction}
In this case by concepts we mean named entities such as names of locations, people and organizations. To recognize these entities multiple learning algorithms or even combinations of learning algorithms have been proposed. The most common are the statistical learning algorithms. Think of algorithms like Maximum Entropy Models, Hidden Markov Models or neural networks. Other proposed algorithms are AdaBoost.MH, employed memory-based learning, transformation-based learning, support vector machines and conditional random fields. All of them in decreasing frequency of use.\cite{EntityRecognition} Florian et al. have shown that combining multiple classifiers can increase the performance, compared to the best performing classifier, by 17-21 \%.\cite{ClassifierCombination} After the entities have been found, labels are assigned to distinguish between a person, an organization or a location. These labels on its own are not worth a lot. They can be used in an application that is build to handle these labels. Therefore 
Abel et al. propose to use the RDF.\cite{AdaptiveSearch} By doing so sources from the Semantic Web (e.g. DBpedia \footnote{http://dbpedia.
org/}) can now be leveraged to enrich the data. Another advantage of using RDF is that RDF is a Semantic Web standard recommended by the W3C\footnote{http://www.w3.org/RDF/}. Because W3C recommends to use this model, most likely a lot of other sources also use this model. This will give a much bigger knowledge base to enrich the data from.

@@

\subsubsection{Emoticons}
Other than semantical analysis we can also use other data from tweets. A good sentiment indicator are emoticons. Twitter users use emoticons (e.g. ":-)", ":(" or "=(") to indicate their mood.\cite{twitterSentiment} Since the number of emoticons is limited a sentiment can easily be assigned manually. The information from the emoticons can be leveraged in two ways
\begin{enumerate}
\item \textbf{Labeling:} The emoticons can be used to assign a sentiment to a tweet. This can be a very useful asset to take into account when assigning a sentiment to a tweet. The emoticons can be considered unigram features for some classifier. There they will be strong evidence for a certain sentiment.
\item \textbf{Training:} Because of their clear evidence for a sentiment emoticons can also be used to build a training corpus for a supervised learning algorithm. Twitter can be queried for "positive" emoticons and "negative" emoticons and then, by assuming the emoticon reflects the sentiment of message, the tweet can be labeled accordingly.\cite{moodLens} 
\end{enumerate}
We now see that emoticons can be used to extract a sentiment from a tweet. However, they are not sufficient to just use emoticons to label the tweets. Simply because not all tweets contain emoticons.
\subsubsection{Usage}
The possibilities of sentiment analysis are not too hard to imagine. Every company wants to know how their newly launched product is doing, film makes want to know what people think about their new movie and politicians want to know how they are doing for the next elections. Using sentiment analysis all this information will become available\cite{machineLearning}.

>>>>

\section{Single tweet analysis}
In this section we will discuss the different types of analysis that can be done on one single tweet. This usually is also the basis for a analysis of a larger amount of tweets or even on a stream of tweets. We will start of by discussing Data enrichment. This part will elaborate on the different techniques to enrich tweets in order to be able to do better analysis on that tweet. After that we will discuss Semantic analysis. In this section an overview is given of different techniques to do a sentiment analysis and how data enrichment can be used to improve the performance.
\subsection{Data enrichment}
At the very basis of the analysis of a tweet is data enrichment. Because the length of a tweet is limited to 140 characters, the users of Twitter use language in creative ways. These ways of communicating makes it sometimes, even for humans, hard to understand what a tweet is about. To get a better understanding of the meaning of a tweet, a context is needed. For humans a context is created by linking the concepts from the tweet to earlier acquired knowledge. The same thing can be done by computers. Two stages can be defined in this process, first the extraction of the concepts from the tweet, after that a context can be created by linking the tweet to other sources.
\subsubsection{Concept extraction}
In this paper when we say concepts we mean named entities such as names of locations, people and organizations. To recognize these entities multiple learning algorithms or even combinations of learning algorithms have been proposed. Nowadays the most common are the statistical learning algorithms. Think of algorithms like Maximum Entropy Models, Hidden Markov Models or neural networks. Other proposed algorithms are AdaBoost.MH, employed memory-based learning, transformation-based learning, support vector machines and conditional random fields. All of them in decreasing frequency of use.\cite{EntityRecognition} Florian et al. have shown that combining multiple classifiers can increase the performance, compared to the best performing classifier, by 17-21 \%.\cite{ClassifierCombination} To do this they set up a framework to compare the results of different (combined) classifiers.

After the entities have been found, labels are assigned to distinguish between a person, an organization or a location. Custom labels are not worth a lot. They can be used in an application that is specially build to handle these labels. To deal with this problem Abel et al. propose to use the RDF (Resource Description Framework).\cite{AdaptiveSearch} By using RDF to represent the data, sources from the Semantic Web (e.g. DBpedia \footnote{http://dbpedia.
org/}) can now be leveraged to enrich the data because RDF is a Semantic Web standard recommended by the W3C\footnote{http://www.w3.org/RDF/}. Because W3C recommends to use this model, most likely a lot of resources also use this model. This will give a much bigger knowledge base to enrich the data from.

@@

\subsubsection{Emoticons}
Other than semantical analysis we can also use other data from tweets. A good sentiment indicator are emoticons. Twitter users use emoticons (e.g. ":-)", ":(" or "=(") to indicate their mood.\cite{twitterSentiment} Since the number of emoticons is limited a sentiment can easily be assigned manually. The information from the emoticons can be leveraged in two ways
\begin{enumerate}
\item \textbf{Labeling:} The emoticons can be used to assign a sentiment to a tweet. This can be a very useful asset to take into account when assigning a sentiment to a tweet. The emoticons can be considered unigram features for some classifier. There they will be strong evidence for a certain sentiment.
\item \textbf{Training:} Because of their clear evidence for a sentiment emoticons can also be used to build a training corpus for a supervised learning algorithm. Twitter can be queried for "positive" emoticons and "negative" emoticons and then, by assuming the emoticon reflects the sentiment of message, the tweet can be labeled accordingly.\cite{moodLens} 
\end{enumerate}
We now see that emoticons can be used to extract a sentiment from a tweet. However, they are not sufficient to just use emoticons to label the tweets. Simply because not all tweets contain emoticons.
\subsubsection{Usage}
The possibilities of sentiment analysis are not too hard to imagine. Every company wants to know how their newly launched product is doing, film makes want to know what people think about their new movie and politicians want to know how they are doing for the next elections. Using sentiment analysis all this information will become available as Pang et al. \cite{machineLearning} have shown in their research. They tried to classify documents based on their sentiment. In their research they used movie reviews as documents. They found that a semantical classifier outperforms human classification but compared to topic-based classification, sentiment classification falls behind. So even though there are enough possible applications for sentiment analysis, there is still a lot of room for improvement in this field.
